### 2025-08-10 base-rates.qmd
20250809 - label functions

The [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] contains the `pAgivenB()` function that estimates the probability of one event, $A$, given another event, $B$.

@@

We can then estimate the marginal probability of the event, substititing in $P(R | \text{not } C)$, using the law of total probability.
The [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] contains the `pA()` function that estimates the marginal probability of one event, $A$.

@@

The [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] contains the `pBgivenNotA()` function that estimates the probability of one event, $B$, given that another event, $A$, did not occur.

@@

The [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] contains the `pB()` function that estimates the marginal probability of one event, $B$.

@@

The [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] contains the `posttestProbability()` function that estimates the posttest probability of an event, given the [pretest probability](#sec-baseRate) and the [likelihood ratio](#sec-diagnosticLikelihoodRatio), or given the [pretest probability](#sec-baseRate) and the [sensitivity](#sec-sensitivity) (SN) and [specificity](#sec-specificity) (SP) of the test.

@@

A probability nomogram calculator is available from @Schwartz2006 at the following link: [http://araw.mede.uic.edu/cgi-bin/testcalc.pl](http://araw.mede.uic.edu/cgi-bin/testcalc.pl) (archived at <https://perma.cc/X8TF-7YBX>).
The [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] contains the `nomogrammer()` function that creates a nomogram plot using the [positive](#sec-positiveLikelihoodRatio) and [negative](#sec-negativeLikelihoodRatio) [likelihood ratio](#sec-diagnosticLikelihoodRatio) or using the [sensitivity](#sec-sensitivity) (SN) and [specificity](#sec-specificity) (SP) of the test, as adapted from @Chekroud2017:

>>>>

The [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] contains the `petersenlab::pAgivenB()` function that estimates the probability of one event, $A$, given another event, $B$.

@@

We can then estimate the marginal probability of the event, substititing in $P(R | \text{not } C)$, using the law of total probability.
The [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] contains the `petersenlab::pA()` function that estimates the marginal probability of one event, $A$.

@@

The [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] contains the `petersenlab::pBgivenNotA()` function that estimates the probability of one event, $B$, given that another event, $A$, did not occur.

@@

The [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] contains the `petersenlab::pB()` function that estimates the marginal probability of one event, $B$.

@@

The [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] contains the `petersenlab::posttestProbability()` function that estimates the posttest probability of an event, given the [pretest probability](#sec-baseRate) and the [likelihood ratio](#sec-diagnosticLikelihoodRatio), or given the [pretest probability](#sec-baseRate) and the [sensitivity](#sec-sensitivity) (SN) and [specificity](#sec-specificity) (SP) of the test.

@@

A probability nomogram calculator is available from @Schwartz2006 at the following link: [http://araw.mede.uic.edu/cgi-bin/testcalc.pl](http://araw.mede.uic.edu/cgi-bin/testcalc.pl) (archived at <https://perma.cc/X8TF-7YBX>).
The [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] contains the `petersenlab::nomogrammer()` function that creates a nomogram plot using the [positive](#sec-positiveLikelihoodRatio) and [negative](#sec-negativeLikelihoodRatio) [likelihood ratio](#sec-diagnosticLikelihoodRatio) or using the [sensitivity](#sec-sensitivity) (SN) and [specificity](#sec-specificity) (SP) of the test, as adapted from @Chekroud2017:

### 2025-08-10 basic-statistics.qmd
20250809 - label functions

The Hodges-Lehmann statistic (aka pseudomedian) is computed as the median of all pairwise means, and it is also robust to outliers.
The pseudomedian can be computed using the `DescTools` package [@R-DescTools].
The pseudomedian of the values of 0, 0, 10, 15, 20, 30, and 1000 is `{r} exampleValues_pseudomedian`.
The pseudomedian is an example of a robust average—an index of the center of a distribution that is less sensitive to outliers.

@@

The mode is the most common/frequent value.
The mode of the values of 0, 0, 10, 15, 20, 30, and 1000 is `{r} exampleValues_mode`.
The [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] contains the `Mode()` function for computing the mode of a set of data.

@@

- mean:

@@

- median:

@@

- Hodges-Lehmann statistic (aka pseudomedian):

@@

- mode:

@@

- weighted mean:

@@

- weighted median:

@@

- standard deviation:

@@

- variance:

@@

- range:

@@

- minimum and maximum:

@@

- quantiles (in this case, the 10th and 90th quantiles):

@@

- weighted quantiles (in this case, the 10th and 90th quantiles):

@@

- interquartile range (IQR):

@@

- median absolute deviation:

@@

- skewness:

@@

- kurtosis:

@@

You can easily convert a variable to a *z* score using the `scale()` function:

@@

The area under a normal curve within one standard deviation of the mean is calculated below using the `pnorm()` function, which calculates the cumulative density function for a normal curve.

@@

If you want to determine the *z* score associated with a particular percentile in a normal distribution, you can use the `qnorm()` function.
For instance, the *z* score associated with the 37th percentile is:

@@

One-tailed test:

@@

Two-tailed test:

@@

One-tailed test:

@@

Two-tailed test:

@@

Factor A (one-tailed test):

@@

Factor B (one-tailed test):

@@

Interaction (one-tailed test):

@@

Factor A (two-tailed test):

@@

Factor B (two-tailed test):

@@

Interaction (two-tailed test):

@@

One-tailed test:

@@

Two-tailed test:

@@

Solving for statistical power achieved (given [effect size](#sec-practicalSignificance), sample size, and [alpha level](#sec-statisticalSignificance)):

@@

Solving for statistical power achieved (given [effect size](#sec-practicalSignificance), sample size per group, and [alpha level](#sec-statisticalSignificance)):

@@

Solving for statistical power achieved (given [effect size](#sec-practicalSignificance), sample size per group, and [alpha level](#sec-statisticalSignificance)):

@@

Solving for statistical power achieved (given [effect size](#sec-practicalSignificance), sample size per group, and [alpha level](#sec-statisticalSignificance)):

@@

@@

@@

```{r}
monteCarloSim_results <- SimDesign::runSimulation(
  design       = Design,
  replications = 1000,
  generate     = Generate,
  analyse      = Analyse,
  summarise    = Summarise,
  seed = genSeeds(Design, iseed = 52242), # for reproducibility
  parallel = TRUE # for faster (parallel) processing
)
```

@@

Solving for statistical power achieved (given [effect size](#sec-practicalSignificance), sample size per group, and [alpha level](#sec-statisticalSignificance)):

@@

Solving for statistical power achieved (given [effect size](#sec-practicalSignificance), sample size per group, and [alpha level](#sec-statisticalSignificance)):

@@

Solving for statistical power achieved (given [effect size](#sec-practicalSignificance), sample size, and [alpha level](#sec-statisticalSignificance)):

@@

Solving for statistical power achieved (given [effect size](#sec-practicalSignificance), sample size, and [alpha level](#sec-statisticalSignificance)):

>>>>

The Hodges-Lehmann statistic (aka pseudomedian) is computed as the median of all pairwise means, and it is also robust to outliers.
The pseudomedian can be computed using the `DescTools::HodgesLehmann()` function of the `DescTools` package [@R-DescTools].
The pseudomedian of the values of 0, 0, 10, 15, 20, 30, and 1000 is `{r} exampleValues_pseudomedian`.
The pseudomedian is an example of a robust average—an index of the center of a distribution that is less sensitive to outliers.

@@

The mode is the most common/frequent value.
The mode of the values of 0, 0, 10, 15, 20, 30, and 1000 is `{r} exampleValues_mode`.
The [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] contains the `petersenlab::Mode()` function for computing the mode of a set of data.

@@

- mean (`base::mean()`):

@@

- median (`stats::median()`):

@@

- Hodges-Lehmann statistic (aka pseudomedian; `DescTools::HodgesLehmann()`):

@@

- mode (`petersenlab::Mode()`):

@@

- weighted mean (`stats::weighted.mean()`):

@@

- weighted median (`petersenlab::wthdquantile()`):

@@

- standard deviation (`stats::sd()`):

@@

- variance (`stats::var()`):

@@

- range (`base::range()`):

@@

- minimum (`base::min()`) and maximum (`base::max()`):

@@

- quantiles (in this case, the 10th and 90th quantiles; `stats::quantile()`):

@@

- weighted quantiles (in this case, the 10th and 90th quantiles; `petersenlab::wthdquantile()`):

@@

- interquartile range (IQR; `stats::IQR()`):

@@

- median absolute deviation (`stats::mad()`):

@@

Below is R code to estimate each using the `psych::skew()` and `psych::kurtosis()` functions of the `psych` package [@R-psych]:

- skewness (`psych::skew()`):

@@

- kurtosis (`psych::kurtosi()`):

@@

You can easily convert a variable to a *z* score using the `base::scale()` function:

@@

The area under a normal curve within one standard deviation of the mean is calculated below using the `stats::pnorm()` function, which calculates the cumulative density function for a normal curve.

@@

If you want to determine the *z* score associated with a particular percentile in a normal distribution, you can use the `stats::qnorm()` function.
For instance, the *z* score associated with the 37th percentile is:

@@

One-tailed test (`stats::qt()`):

@@

Two-tailed test (`stats::qt()`):

@@

One-tailed test (`stats::qf()`):

@@

Two-tailed test (`stats::qf()`):

@@

Factor A (one-tailed test; `stats::qf()`):

@@

Factor B (one-tailed test; `stats::qf()`):

@@

Interaction (one-tailed test; `stats::qf()`):

@@

Factor A (two-tailed test; `stats::qf()`):

@@

Factor B (two-tailed test; `stats::qf()`):

@@

Interaction (two-tailed test; `stats::qf()`):

@@

One-tailed test (`stats::qchisq()`):

@@

Two-tailed test (`stats::qchisq()`):

@@

Solving for statistical power achieved (given [effect size](#sec-practicalSignificance), sample size, and [alpha level](#sec-statisticalSignificance); `pwr::pwr.t.test()`):

@@

Solving for statistical power achieved (given [effect size](#sec-practicalSignificance), sample size per group, and [alpha level](#sec-statisticalSignificance); `pwr::pwr.t.test()`):

@@

Solving for statistical power achieved (given [effect size](#sec-practicalSignificance), sample size per group, and [alpha level](#sec-statisticalSignificance); `pwr::pwr.t2n.test()`):

@@

Solving for statistical power achieved (given [effect size](#sec-practicalSignificance), sample size per group, and [alpha level](#sec-statisticalSignificance); `pwr::pwr.anova.test()`):

@@

Solving for sample size per group needed (given [effect size](#sec-practicalSignificance), power, and [alpha level](#sec-statisticalSignificance); `pwr::pwr.anova.test()`):

@@

Solving for the minimum detectable [effect size](#sec-practicalSignificance) (given sample size per group, power, and [alpha level](#sec-statisticalSignificance); `pwr::pwr.anova.test()`):

@@

```{r}
#| output: false

monteCarloSim_results <- SimDesign::runSimulation(
  design       = Design,
  replications = 1000,
  generate     = Generate,
  analyse      = Analyse,
  summarise    = Summarise,
  seed = genSeeds(Design, iseed = 52242), # for reproducibility
  parallel = TRUE # for faster (parallel) processing
)
```

@@

Solving for statistical power achieved (given [effect size](#sec-practicalSignificance), sample size per group, and [alpha level](#sec-statisticalSignificance); `WebPower::wp.rmanova()`):

@@

Solving for statistical power achieved (given [effect size](#sec-practicalSignificance), sample size per group, and [alpha level](#sec-statisticalSignificance); `pwr::pwr.r.test()`):

@@

Solving for statistical power achieved (given [effect size](#sec-practicalSignificance), sample size, and [alpha level](#sec-statisticalSignificance); `pwr::pwr.f2.test()`):

@@

Solving for statistical power achieved (given [effect size](#sec-practicalSignificance), sample size, and [alpha level](#sec-statisticalSignificance); `pwr::pwr.chisq.test()`):

### 2025-08-10 correlation.qmd
20250802 - documentation
20250809 - label functions

A partial correlation examines the association between variables after controlling for one or more variables.
Below, we examine a partial [Pearson correlation](#sec-correlationExamplesPearson) between height and fantasy points after controlling for age using the the `partial.r()` function from the `psych` package [@R-psych].

```{r}
psych::partial.r(
  data = player_stats_seasonal,
  x = c("height","fantasyPoints"),
  y = "age"
)
```

@@

An example of a nonlinear correlation is a distance correlation and the xi ($\xi$) coefficient.
These estimates of nonlinear correlation only range from 0–1 (where 0 represents no association and 1 represents a strong association), so they are interpreted differently than a traditional correlation coefficient (which can be negative).
We can estimate a distance correlation using the `correlation()` function of the correlation package [@R-correlation; @Makowski2020_packages].

@@

The [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] contains the `cor.table()` function that generates a correlation matrix of variables, including the $r$ value, number of observations ($n$), asterisks denoting statistical significance, and $p$-value for each association.
The asterisks follow the following traditional convention for statistical significance: $^\dagger p < .1$; $^* p < .05$; $^{**} p < .01$; $^{***} p < .001$.
Here is a [Pearson correlation](#sec-correlationExamplesPearson) matrix:

@@

The [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] contains the `partialcor.table()` function that generates a [partial correlation](#sec-correlationExamplesPartial) matrix.
Here is a [partial correlation](#sec-correlationExamplesPartial) matrix, controlling for the player's age:

@@

We can depict a correlation matrix plot (correlogram) using the `ggcorr()` function of the `GGally` [@R-GGally] package, as shown in @fig-correlogram.

@@

One approach is the biweight midcorrelation, which is based on the median rather than the mean, and is thus less sensitive to outliers.
Another approach is the percentage bend correlation, which gives less weight to observations that are farther away from the median.
We can estimate the each using the `correlation()` function of the correlation package [@R-correlation; @Makowski2020_packages].

>>>>

A partial correlation examines the association between variables after controlling for one or more variables.
Below, we examine a partial [Pearson correlation](#sec-correlationExamplesPearson) between height and fantasy points after controlling for age using the the `psych::partial.r()` function from the `psych` package [@R-psych].

```{r}
psych::partial.r(
  data = player_stats_seasonal,
  x = c("height","fantasyPoints"),
  y = "age" # control variable(s)
)
```

@@

An example of a nonlinear correlation is a distance correlation and the xi ($\xi$) coefficient.
These estimates of nonlinear correlation only range from 0–1 (where 0 represents no association and 1 represents a strong association), so they are interpreted differently than a traditional correlation coefficient (which can be negative).
We can estimate a distance correlation using the `correlation::correlation()` function of the correlation package [@R-correlation; @Makowski2020_packages].

@@

The [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] contains the `petersenlab::cor.table()` function that generates a correlation matrix of variables, including the $r$ value, number of observations ($n$), asterisks denoting statistical significance, and $p$-value for each association.
The asterisks follow the following traditional convention for statistical significance: $^\dagger p < .1$; $^* p < .05$; $^{**} p < .01$; $^{***} p < .001$.
Here is a [Pearson correlation](#sec-correlationExamplesPearson) matrix:

@@

The [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] contains the `petersenlab::partialcor.table()` function that generates a [partial correlation](#sec-correlationExamplesPartial) matrix.
Here is a [partial correlation](#sec-correlationExamplesPartial) matrix, controlling for the player's age:

@@

We can depict a correlation matrix plot (correlogram) using the `GGally::ggcorr()` function of the `GGally` [@R-GGally] package, as shown in @fig-correlogram.

@@

One approach is the biweight midcorrelation, which is based on the median rather than the mean, and is thus less sensitive to outliers.
Another approach is the percentage bend correlation, which gives less weight to observations that are farther away from the median.
We can estimate the each using the `correlation::correlation()` function of the correlation package [@R-correlation; @Makowski2020_packages].

### 2025-08-10 data-visualization.qmd
20250806 - duplicate identifiers
20250809 - label functions

It is also important to use color schemes with distinguishable colors.
Good color schemes for sequential, diverging, and qualitative (i.e., categorical) data are provided by ColorBrewer (<https://colorbrewer2.org>) and are available using the `scale_color_brewer()` and `scale_fill_brewer()` functions of the `ggplot2` package [@R-ggplot2], as demonstrated in @fig-barPlotColorScheme (@sec-barPlotModifiedColorScheme).
Good colorblind-friendly color schemes for binned (`_b`), continuous (`_c`), and discrete (i.e., categorical; `_d`) scales are available using the `scale_color_viridis_X()` and `scale_fill_viridis_X()` functions of the `ggplot2` package [@R-ggplot2] (where `X` is `b`, `c`, or `d`, depending on whether you want a binned, continuous, discrete scale). 
There are a variety of resources for color schemes that are accessible to color-blind viewers:

@@

The `coord_cartesian()` function can be used to specify the axis limits; however, this function is not required and the plot will apply sensible limits by default.
In addition, there are optional settings you can specify or modify, such as [axis titles](#sec-scatterplotPlotLabels), [axis labels](#sec-scatterplotAxes), [faceting](#sec-faceting), and the [theme](#sec-scatterplotTheme).

@@

You can see a full list of color names available using the `colors()` function:

@@

You can use any color by specifying its values in red-green-blue (RGB) color space using the `rgb()` function, as in @fig-colorByRGB.

@@

We can specify multiple colors manually using the `scale_fill_manual()` function, as in @fig-multipleColorsManually.

@@

We can also specify the colors using a palette, such as by the ColorBrewer palette using the `scale_fill_brewer()` function (as in @fig-multipleColorsByPaletteColorBrewer) or the viridis palette using the `scale_fill_viridis_d()` function (as in @fig-multipleColorsByPaletteViridis).

@@

We can specify the line types manually using the `scale_linetype_manual()` function as in @fig-lineTypeByGroupManually.

@@

### Line Width {#sec-ggplot2SyntaxLineType}

@@

We can specify the point shape manually using the `scale_shape_manual()` function as in @fig-pointShapeByGroupManually.

@@

Second, we create the base layer of the plot using the `ggplot()` function of the `ggplot2` package [@R-ggplot2], as in @fig-scatterplot1.
We specify the data object and the variables in the data object that are associated with the x- and y-axes:

@@

Third, we create a scatterplot using the `geom_point()` function from the `ggplot2` package [@R-ggplot2], as in @fig-scatterplot2:

@@

After creating our plot, we can make the plot interactive using the `ggplotly()` function from the `plotly` package [@R-plotly; @Sievert2020_packages], as in @fig-scatterplot11.

@@

After creating our plot, we can make the plot interactive using the `ggplotly()` function from the `plotly` package [@R-plotly; @Sievert2020_packages].

>>>>

It is also important to use color schemes with distinguishable colors.
Good color schemes for sequential, diverging, and qualitative (i.e., categorical) data are provided by ColorBrewer (<https://colorbrewer2.org>) and are available using the `ggplot2::scale_color_brewer()` and `ggplot2::scale_fill_brewer()` functions of the `ggplot2` package [@R-ggplot2], as demonstrated in @fig-barPlotColorScheme (@sec-barPlotModifiedColorScheme).
Good colorblind-friendly color schemes for binned (`_b`), continuous (`_c`), and discrete (i.e., categorical; `_d`) scales are available using the `ggplot2::scale_color_viridis_X()` and `ggplot2::scale_fill_viridis_X()` functions of the `ggplot2` package [@R-ggplot2] (where `X` is `b`, `c`, or `d`, depending on whether you want a binned, continuous, discrete scale). 
There are a variety of resources for color schemes that are accessible to color-blind viewers:

@@

The `ggplot2::coord_cartesian()` function can be used to specify the axis limits; however, this function is not required and the plot will apply sensible limits by default.
In addition, there are optional settings you can specify or modify, such as [axis titles](#sec-scatterplotPlotLabels), [axis labels](#sec-scatterplotAxes), [faceting](#sec-faceting), and the [theme](#sec-scatterplotTheme).

@@

You can see a full list of color names available using the `grDevices::colors()` function:

@@

You can use any color by specifying its values in red-green-blue (RGB) color space using the `grDevices::rgb()` function, as in @fig-colorByRGB.

@@

We can specify multiple colors manually using the `ggplot2::scale_fill_manual()` function, as in @fig-multipleColorsManually.

@@

We can also specify the colors using a palette, such as by the ColorBrewer palette using the `ggplot2::scale_fill_brewer()` function (as in @fig-multipleColorsByPaletteColorBrewer) or the viridis palette using the `ggplot2::scale_fill_viridis_d()` function (as in @fig-multipleColorsByPaletteViridis).

@@

We can specify the line types manually using the `ggplot2::scale_linetype_manual()` function as in @fig-lineTypeByGroupManually.

@@

### Line Width {#sec-ggplot2SyntaxLineWidth}

@@

We can specify the point shape manually using the `ggplot2::scale_shape_manual()` function as in @fig-pointShapeByGroupManually.

@@

Second, we create the base layer of the plot using the `ggplot2::ggplot()` function of the `ggplot2` package [@R-ggplot2], as in @fig-scatterplot1.
We specify the data object and the variables in the data object that are associated with the x- and y-axes:

@@

Third, we create a scatterplot using the `ggplot2::geom_point()` function from the `ggplot2` package [@R-ggplot2], as in @fig-scatterplot2:

@@

After creating our plot, we can make the plot interactive using the `plotly::ggplotly()` function from the `plotly` package [@R-plotly; @Sievert2020_packages], as in @fig-scatterplot11.

@@

After creating our plot, we can make the plot interactive using the `plotly::ggplotly()` function from the `plotly` package [@R-plotly; @Sievert2020_packages].

### 2025-08-10 decision-making.qmd
20250809 - label functions

The [`petersenlab` package](https://cran.r-project.org/web/packages/petersenlab/index.html) [@R-petersenlab] has the `wisdomOfCrowd()` function that computes the overall accuracy of the crowd-averaged projections, including the bracketing rate of the individual projections.

>>>>

The [`petersenlab` package](https://cran.r-project.org/web/packages/petersenlab/index.html) [@R-petersenlab] has the `petersenlab::wisdomOfCrowd()` function that computes the overall accuracy of the crowd-averaged projections, including the bracketing rate of the individual projections.

### 2025-08-10 download-football-data.qmd
20250730 - updates
20250809 - label functions

To download play-by-play data from prior weeks and seasons, we can use the `load_pbp()` function of the `nflreadr` package [@R-nflreadr].
We add a progress bar using the `with_progress()` function from the `progressr` package [@R-progressr] because it takes a while to run.
@HoPBP provide a Data Dictionary for the play-by-play data at the following link: <https://nflreadr.nflverse.com/articles/dictionary_pbp.html>

@@

We can download historical week-by-week actual player statistics using the `load_player_stats()` function from the `nflreadr` package [@R-nflreadr].
@HoPlayerStats provide a Data Dictionary for statistics for offensive players at the following link: <https://nflreadr.nflverse.com/articles/dictionary_player_stats.html>.
@HoPlayerStatsDefense provide a Data Dictionary for statistics for defensive players at the following link: <https://nflreadr.nflverse.com/articles/dictionary_player_stats_def.html>.

@@

A Data Dictionary for the variables is available in the `nfl_stats_variables` object that is returned when running the `calculate_stats()` function:

@@

Second, we can compute the players' career statistics using the `calculate_stats()` function from the `nflfastR` package [@R-nflfastR].

@@

We already load players' week-by-week statistics [above](#sec-downloadActualStats) using the `load_player_stats()` function from the `nflreadr` [@R-nflreadr] package.
Nevertheless, we could compute players' weekly statistics from the play-by-play data using the `calculate_stats()` function from the `nflfastR` [@R-nflfastR] package:

>>>>

To download play-by-play data from prior weeks and seasons, we can use the `nflreadr::load_pbp()` function of the `nflreadr` package [@R-nflreadr].
We add a progress bar using the `progressr::with_progress()` function from the `progressr` package [@R-progressr] because it takes a while to run.
@HoPBP provide a Data Dictionary for the play-by-play data at the following link: <https://nflreadr.nflverse.com/articles/dictionary_pbp.html>

@@

We can download historical week-by-week actual player statistics using the `nflreadr::load_player_stats()` function from the `nflreadr` package [@R-nflreadr].
@HoPlayerStats provide a Data Dictionary for statistics for offensive players at the following link: <https://nflreadr.nflverse.com/articles/dictionary_player_stats.html>.
@HoPlayerStatsDefense provide a Data Dictionary for statistics for defensive players at the following link: <https://nflreadr.nflverse.com/articles/dictionary_player_stats_def.html>.

@@

A Data Dictionary for the variables is available in the `nfl_stats_variables` object that is returned when running the `nflfastR::calculate_stats()` function:

@@

Second, we can compute the players' career statistics using the `nflfastR::calculate_stats()` function from the `nflfastR` package [@R-nflfastR].

@@

We already load players' week-by-week statistics [above](#sec-downloadActualStats) using the `nflreadr::load_player_stats()` function from the `nflreadr` [@R-nflreadr] package.
Nevertheless, we could compute players' weekly statistics from the play-by-play data using the `nflfastR::calculate_stats()` function from the `nflfastR` [@R-nflfastR] package:

### 2025-08-10 evaluating-prediction-accuracy.qmd
20250809 - lineup optimization
20250809 - label functions

The [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] contains the `accuracyAtCutoff()` function that computes many accuracy indices for the prediction of categorical outcome variables.

@@

The [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] contains the `accuracyAtEachCutoff()` function that computes many accuracy indices for the prediction of categorical outcome variables at each possible cutoff.

@@

This section describes threshold-independent indexes of accuracy.
That is, each index of accuracy described in this section provides a single numerical index of accuracy that aggregates the accuracy across all possible cutoffs.
The [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] contains the `accuracyOverall()` function that computes many threshold-independent accuracy indices.

@@

Note: MPE is undefined when one or more of the observed values equals zero, due to division by zero.
The `accuracyOverall()` function of the [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] provides the option in the function to drop undefined values so you can still generate an estimate of accuracy despite undefined values.

@@

Note: MAPE is undefined when one or more of the observed values equals zero, due to division by zero.
The `accuracyOverall()` function of the [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] provides the option in the function to drop undefined values so you can still generate an estimate of accuracy despite undefined values.

@@

Note: sMAPE is undefined when one or more of the individuals has a prediction–observed combination such that the sum of the absolute value of the predicted value and the absolute value of the observed value equals zero ($|\text{predicted}_i| + |\text{observed}_i|$), due to division by zero.
The `accuracyOverall()` function of the [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] provides the option in the function to drop undefined values so you can still generate an estimate of accuracy despite undefined values.

@@

Note: Root mean squared log error is undefined when one or more predicted values or actual values equals −1.
When predicted or actual values are -1, this leads to $log(0)$, which is undefined.
The `accuracyOverall()` function of the [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] provides the option in the function to drop undefined values so you can still generate an estimate of accuracy despite undefined values.

@@

The effect size of a predictor, i.e., the standardized regression coefficient is called a beta ($\beta$) coefficient, is a general index of [discrimination accuracy](#sec-discrimination) for a continuous outcome.
Larger values reflect greater accuracy.
We can obtain standardized regression coefficients by standardizing the predictors and outcome using the `scale()` function in `R`.

@@

Below are estimates of projection accuracy and [calibration plots](#sec-calibrationPlot) for [seasonal](#sec-accuracyProjectionsSeasonal) and [weekly](#sec-accuracyProjectionsWeekly) projections.
FantasyFootballAnalytics.net also provides tools to evaluate the accuracy of historical projections [@FFA-accuracy]: <https://fantasyfootballanalytics.net/ffanalytics-web-app-accuracy-tab>.

@@

A calibration plot of the seasonal projections for all players (across all positions) is in @fig-calibrationPlotSeasonalOverall.
The best-fit LOESS line is the yellow line.
The intercept from the best-fit line is positive; the slope is less than one.
The predicted values are less than the observed values at lower levels of projected points, whereas the predicted values are higher than the observed values at higher levels of projected points, consistent with the overextremity form of [miscalibration](#sec-calibration).
The confidence interval of the observed value (i.e., the purple band) is the interval within which we have 95% confidence that the true observed value would lie for a given predicted value, based on the model.
The 95% prediction interval of the observed value (i.e., the dashed red lines) is the interval within which we would expect that 95% of future observations would lie for a given predicted value.
We calculate the 95% prediction interval based on a LOESS model, which allows nonlinear associations, using the `loess.sd()` function of the `msir` package [@R-msir; @Scrucca2011_packages].
The black diagonal line indicates the reference line with an intercept of zero and a slope of one.
The predictions would be significantly [miscalibrated](#sec-calibration) at a given level of the predicted values if the 95% confidence interval of the observed value does not include the reference line at that level of the predicted value.
In this case, based on the intercept and slope of the [calibration plot](#sec-calibrationPlot) in @fig-calibrationPlotSeasonalOverall, the 95% confidence interval of the observed value does not include the reference line (i.e., the actual observed value) at lower or upper levels of the predicted values, so the predictions are [miscalibrated](#sec-calibration) at lower and upper levels of the predicted values.

>>>>

The [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] contains the `petersenlab::accuracyAtCutoff()` function that computes many accuracy indices for the prediction of categorical outcome variables.

@@

The [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] contains the `petersenlab::accuracyAtEachCutoff()` function that computes many accuracy indices for the prediction of categorical outcome variables at each possible cutoff.

@@

This section describes threshold-independent indexes of accuracy.
That is, each index of accuracy described in this section provides a single numerical index of accuracy that aggregates the accuracy across all possible cutoffs.
The [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] contains the `petersenlab::accuracyOverall()` function that computes many threshold-independent accuracy indices.

@@

Note: MPE is undefined when one or more of the observed values equals zero, due to division by zero.
The `petersenlab::accuracyOverall()` function of the [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] provides the option in the function to drop undefined values so you can still generate an estimate of accuracy despite undefined values.

@@

Note: MAPE is undefined when one or more of the observed values equals zero, due to division by zero.
The `petersenlab::accuracyOverall()` function of the [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] provides the option in the function to drop undefined values so you can still generate an estimate of accuracy despite undefined values.

@@

Note: sMAPE is undefined when one or more of the individuals has a prediction–observed combination such that the sum of the absolute value of the predicted value and the absolute value of the observed value equals zero ($|\text{predicted}_i| + |\text{observed}_i|$), due to division by zero.
The `petersenlab::accuracyOverall()` function of the [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] provides the option in the function to drop undefined values so you can still generate an estimate of accuracy despite undefined values.

@@

Note: Root mean squared log error is undefined when one or more predicted values or actual values equals −1.
When predicted or actual values are -1, this leads to $log(0)$, which is undefined.
The `petersenlab::accuracyOverall()` function of the [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] provides the option in the function to drop undefined values so you can still generate an estimate of accuracy despite undefined values.

@@

The effect size of a predictor, i.e., the standardized regression coefficient is called a beta ($\beta$) coefficient, is a general index of [discrimination accuracy](#sec-discrimination) for a continuous outcome.
Larger values reflect greater accuracy.
We can obtain standardized regression coefficients by standardizing the predictors and outcome using the `base::scale()` function in `R`.

@@

Below are estimates of projection accuracy and [calibration plots](#sec-calibrationPlot) for [seasonal](#sec-accuracyProjectionsSeasonal) and [weekly](#sec-accuracyProjectionsWeekly) projections.
FantasyFootballAnalytics.net also provides tools to evaluate the accuracy of historical projections [@FFA-accuracy]: <https://fantasyfootballanalytics.net/ffanalytics-web-app-accuracy-tab>.
For instance, the web application allows evaluating the accuracy of historical projections:
<https://apps.fantasyfootballanalytics.net>.

@@

A calibration plot of the seasonal projections for all players (across all positions) is in @fig-calibrationPlotSeasonalOverall.
The best-fit LOESS line is the yellow line.
The intercept from the best-fit line is positive; the slope is less than one.
The predicted values are less than the observed values at lower levels of projected points, whereas the predicted values are higher than the observed values at higher levels of projected points, consistent with the overextremity form of [miscalibration](#sec-calibration).
The confidence interval of the observed value (i.e., the purple band) is the interval within which we have 95% confidence that the true observed value would lie for a given predicted value, based on the model.
The 95% prediction interval of the observed value (i.e., the dashed red lines) is the interval within which we would expect that 95% of future observations would lie for a given predicted value.
We calculate the 95% prediction interval based on a LOESS model, which allows nonlinear associations, using the `msir::loess.sd()` function of the `msir` package [@R-msir; @Scrucca2011_packages].
The black diagonal line indicates the reference line with an intercept of zero and a slope of one.
The predictions would be significantly [miscalibrated](#sec-calibration) at a given level of the predicted values if the 95% confidence interval of the observed value does not include the reference line at that level of the predicted value.
In this case, based on the intercept and slope of the [calibration plot](#sec-calibrationPlot) in @fig-calibrationPlotSeasonalOverall, the 95% confidence interval of the observed value does not include the reference line (i.e., the actual observed value) at lower or upper levels of the predicted values, so the predictions are [miscalibrated](#sec-calibration) at lower and upper levels of the predicted values.
